Using Incremental Many-to-One Queries to Build a Fast and Tight Heuristic for A* in Road Networks

### 2022-08-30 paper/ch_potentials.tex
augmented weights

\begin{algorithm2e}[tb]
\KwData{$D^{\downarrow}[v]$: tentative distance from any vertex $v \in V$ to target $t$ in $\gchd$}
\KwData{Minimum priority queue $Q$, ordered by tentative distances}
$D^{\downarrow}[v] \gets +\infty$ for all $v\neq t$\;
$D^{\downarrow}[t] \gets 0$\;
Make $Q$ only contain $t$ with tentative distance $0$\;
\While{$Q$ not empty}{
  $u\gets$ pop minimum element from $Q$\;
  \For{all reversed downward edges $uv \in \rechd$}{
    \If{$D^{\downarrow}[u] + \overleftarrow{w}(uv) < D^{\downarrow}[v]$}{
      $D^{\downarrow}[v] \gets D^{\downarrow}[u] + \overleftarrow{w}(uv)$\;
      Add $v$ or decrease $v$'s key in $Q$ to $D^{\downarrow}[v]$\;
    }
  }
}
\caption{CH backward search.}
\label{algo:ch-backward}
\end{algorithm2e}

\emph{Contraction Hierarchies} (CH) is a two-phase speedup technique to accelerate shortest path computations on road networks through precomputation.
For a detailed discussion we refer to~\cite{gssv-erlrn-12}.
Here, we only briefly introduce the necessary notation and algorithms used in this paper.
In a preprocessing phase, vertices are ordered totally by ``importance'' where more important vertices should lie on more shortest paths.
Intuitively, vertices on highways are more important than vertices on rural streets.
For CH, such an ordering is obtained heuristically.
The position of a vertex in the order is also denoted as its \emph{rank}.
Vertices of higher rank are informally referred to and visualized as ``higher up'' in the hierarchy.
Therefore, an edge where the tail has a lower rank than the head is an \emph{upward} edge.
Analogously, when the head vertex has the lower rank, the edge is said to go \emph{downward}.
Once such an importance order was obtained, all vertices are contracted successively by ascending importance.
To \emph{contract} a vertex means temporarily removing it from the graph while inserting \emph{shortcut edges} between more important neighbours to preserve shortest distances among them.
The result is an \emph{augmented graph} $G^+$ with original edges and shortcuts.
We often refer to the augmented graph split into an upward graph $\gchu = (V, \echu)$ containing only upward edges and a downward graph $\gchd = (V, \echu)$ containing only downward edges.

@@

\begin{algorithm2e}[tb]
\KwData{$D[v]$: tentative distance from any $v \in V$ to $t$}
Execute Algorithm~\ref{algo:ch-backward}, filling $D$\;
\For{all CH levels $L$ from most to least important}{
	\For{all up-edges $uv \in \echu$ with $u$ in $L$}{
		\If{$D[u] > D[v] + w(uv)$}{
			$D[u] \leftarrow D[v] + w(uv)$\;
		}
	}
}
\caption{PHAST basic all-to-one search.}
\label{algo:phast}
\end{algorithm2e}

@@

\Fn{\Dist{$u$}}{
	\If{$D[u] = \bot$}{
		$D[u]\leftarrow D^{\downarrow}[u]$\;
		\For{all up-edges $uv \in \echu$}{
      $D[u]\leftarrow\min(D[u],w(uv)+\Dist(v))$\;
		}
	}
	\Return{$D[u]$}\;
}
\caption{Lazy RPHAST algorithm.}
\label{algo:lazy_rphast_ch}
\end{algorithm2e}

@@

\Fn{\Dist{$u$}}{
  \tcp{Determine the vertices $v$ for which $D[v]$ needs to be computed}
  $v \gets u$\;
	\While{$D[v] = \bot$}{
    Push $v$ onto $S$\;
		\If{$P[v] = \bot$}{
      \Break\;
		}
    $v \gets P[v]$\;
  }
  \tcp{Compute $D$ for those vertices}
  \While{$S$ not empty}{
    $v\leftarrow$ pop top element from $S$\;
		$D[v]\gets D^{\downarrow}[v]$\;
		\For{all up-edges $vw \in \echu$}{
      $D[v]\gets\min(D[v],w(vw)+D[w])$\;
		}
	}
	\Return{$D[u]$}\;
}
\caption{Elimination tree-based Lazy RPHAST algorithm.}
\label{algo:cch_pot}
\end{algorithm2e}

>>>>

\begin{algorithm2e}[tb]
\KwData{$D^{\downarrow}[v]$: tentative distance from any vertex $v \in V$ to target $t$ in $\gchd$}
\KwData{Minimum priority queue $Q$, ordered by tentative distances}
$D^{\downarrow}[v] \gets +\infty$ for all $v\neq t$\;
$D^{\downarrow}[t] \gets 0$\;
Make $Q$ only contain $t$ with tentative distance $0$\;
\While{$Q$ not empty}{
  $u\gets$ pop minimum element from $Q$\;
  \For{all reversed downward edges $uv \in \rechd$}{
    \If{$D^{\downarrow}[u] + \overleftarrow{w}^+(uv) < D^{\downarrow}[v]$}{
      $D^{\downarrow}[v] \gets D^{\downarrow}[u] + \overleftarrow{w}^+(uv)$\;
      Add $v$ or decrease $v$'s key in $Q$ to $D^{\downarrow}[v]$\;
    }
  }
}
\caption{CH backward search.}
\label{algo:ch-backward}
\end{algorithm2e}

\emph{Contraction Hierarchies} (CH) is a two-phase speedup technique to accelerate shortest path computations on road networks through precomputation.
For a detailed discussion we refer to~\cite{gssv-erlrn-12}.
Here, we only briefly introduce the necessary notation and algorithms used in this paper.
In a preprocessing phase, vertices are ordered totally by ``importance'' where more important vertices should lie on more shortest paths.
Intuitively, vertices on highways are more important than vertices on rural streets.
For CH, such an ordering is obtained heuristically.
The position of a vertex in the order is also denoted as its \emph{rank}.
Vertices of higher rank are informally referred to and visualized as ``higher up'' in the hierarchy.
Therefore, an edge where the tail has a lower rank than the head is an \emph{upward} edge.
Analogously, when the head vertex has the lower rank, the edge is said to go \emph{downward}.
Once such an importance order was obtained, all vertices are contracted successively by ascending importance.
To \emph{contract} a vertex means temporarily removing it from the graph while inserting \emph{shortcut edges} between more important neighbours to preserve shortest distances among them.
The result is an \emph{augmented graph} $G^+$ of original edges and shortcuts with weights $w^+$.
We often refer to the augmented graph split into an upward graph $\gchu = (V, \echu)$ containing only upward edges and a downward graph $\gchd = (V, \echu)$ containing only downward edges.

@@

\begin{algorithm2e}[tb]
\KwData{$D[v]$: tentative distance from any $v \in V$ to $t$}
Execute Algorithm~\ref{algo:ch-backward}, filling $D$\;
\For{all CH levels $L$ from most to least important}{
	\For{all up-edges $uv \in \echu$ with $u$ in $L$}{
		\If{$D[u] > D[v] + w^+(uv)$}{
			$D[u] \leftarrow D[v] + w^+(uv)$\;
		}
	}
}
\caption{PHAST basic all-to-one search.}
\label{algo:phast}
\end{algorithm2e}

@@

\Fn{\Dist{$u$}}{
	\If{$D[u] = \bot$}{
		$D[u]\leftarrow D^{\downarrow}[u]$\;
		\For{all up-edges $uv \in \echu$}{
      $D[u]\leftarrow\min(D[u],w^+(uv)+\Dist(v))$\;
		}
	}
	\Return{$D[u]$}\;
}
\caption{Lazy RPHAST algorithm.}
\label{algo:lazy_rphast_ch}
\end{algorithm2e}

@@

\Fn{\Dist{$u$}}{
  \tcp{Determine the vertices $v$ for which $D[v]$ needs to be computed}
  $v \gets u$\;
	\While{$D[v] = \bot$}{
    Push $v$ onto $S$\;
		\If{$P[v] = \bot$}{
      \Break\;
		}
    $v \gets P[v]$\;
  }
  \tcp{Compute $D$ for those vertices}
  \While{$S$ not empty}{
    $v\leftarrow$ pop top element from $S$\;
		$D[v]\gets D^{\downarrow}[v]$\;
		\For{all up-edges $vx \in \echu$}{
      $D[v]\gets\min(D[v],w^+(vx)+D[x])$\;
		}
	}
	\Return{$D[u]$}\;
}
\caption{Elimination tree-based Lazy RPHAST algorithm.}
\label{algo:cch_pot}
\end{algorithm2e}
